---
title: 'EDS 213: Week 2 Assignment'
author: "Paloma Cartwright"
date: "06/10/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(dataRetrieval)
library(metajam)
library(lubridate)
```

# Using dataRetrieval USGS package 

### Ventura Example from Class

```{r}
siteNumber <- "11118500"
parameter <- "00060"  # Discharge
startDate <- "2019-10-01"  
endDate <- "2020-10-05" 
discharge <- readNWISdv(siteNumber, 
                        parameter, 
                        startDate, 
                        endDate) %>% 
  rename(discharge = X_00060_00003, 
         date = Date) 
ggplot(discharge, aes(x = date, y = discharge)) +
  geom_line()
```


#### If you didn't pass a date argument, what happens?

```{r}
siteNumber2 <- "11118500"
parameter2 <- "00060"  # Discharge
startDate2 <- "" 
endDate2 <- "2020-10-05"
discharge_no_date <- readNWISdv(siteNumber2, parameter2, startDate2, endDate2) %>% 
  rename(date = Date,
         discharge = X_00060_00003) %>% 
  filter(lubridate::year(date) > 1930, 
         year(date) < 1955)
ggplot(discharge_no_date, aes(x = date, y = discharge)) +
  geom_line()
```


## Assignment 2.1: Santa Paula Creek & Ventura River Flows 

```{r}
siteNumberSP <- c("11113500", "11118500")
parameterSP <- "00060" #Discharge 
startDateSP <- "2021-10-01"
endDateSP <- ""

both_info <- readNWISdv(siteNumberSP, parameterSP, startDateSP, endDateSP) %>% 
  rename(date = Date,
         discharge = X_00060_00003)

ggplot(data = both_info, aes(x = date, y = discharge)) +
  geom_line(aes(color = site_no)) +
  labs(y = "Discharge (cubic feet)", 
       x = "Date") +
  scale_color_manual(name = "Site ID", 
                     labels = c("Santa Paula", "Ventura"), 
                     values = c("cyan3", "red"))



```

The best way to query multiple datasets using this package is to pass a vector of site numbers as the first argument to `readNWISdv()`. You can then wrangle the data however you see fit after it is all read in, filtering based on the siteno. 



# Assignment 2.2: Using metajam

```{r, message=FALSE}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"
#metajam::download_d1_data(data_url = data_url, here::here("data"))
data <- read_d1_files(here::here("data", "household_language_files"))
household_data <- data$data
only_english <- household_data %>% 
  clean_names() %>% 
  filter(year >= 2009, 
         year <= 2015, 
         !total == 0) %>% 
  select(year, sasap_region, city, total, speak_only_english) %>% 
  group_by(year) %>% 
  summarize(yearly_percent_avg=mean(speak_only_english/total))
ggplot(only_english, aes(x = year, y = yearly_percent_avg)) +
  geom_line()
```

