---
title: 'EDS 213: Week 2 Assignment'
author: "Paloma Cartwright"
date: "06/10/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(dataRetrieval)
library(metajam)
library(lubridate)
library(scales)
```

# Using dataRetrieval USGS package 

### Ventura Example from Class

```{r}
siteNumber <- "11118500"
parameter <- "00060"  # Discharge
startDate <- "2019-10-01"  
endDate <- "2021-10-01" 
discharge <- readNWISdv(siteNumber, 
                        parameter, 
                        startDate, 
                        endDate) %>% 
  rename(discharge = X_00060_00003, 
         date = Date) 
ggplot(discharge, aes(x = date, y = discharge)) +
  geom_line() +
  labs(x = "Date", 
       y = "Discharge (cubic feet)", 
       title = "Ventura River Discharge (2019-10-01 to 2021-10-01)") +
  theme_bw()
```


#### If you didn't pass a date argument, what happens?

```{r}
siteNumber2 <- "11118500"
parameter2 <- "00060"  # Discharge
startDate2 <- "" 
endDate2 <- "2020-10-05"
discharge_no_date <- readNWISdv(siteNumber2, parameter2, startDate2, endDate2) %>% 
  rename(date = Date,
         discharge = X_00060_00003) %>% 
  filter(lubridate::year(date) > 1930, 
         year(date) < 1970)

ggplot(discharge_no_date, aes(x = as.Date(date), y = discharge)) +
  geom_line() +
  geom_vline(xintercept = as.numeric(as.Date("1947-01-01")),
             color = "red",
             lwd = 1, 
             linetype = 3) +
  labs(x = "Date", 
       y = "Discharge (cubic feet)", 
       title = "Ventura River Discharge (1930 - 1970)")
```

Red line on the graph above shows when the Ojai Dam was built and the impact it had on the discharge of the river immediately after. The discharge was not consistent again in the years following the implementation of this dam.

## Assignment 2.1: Santa Paula Creek & Ventura River Flows 

```{r, fig.width=9, fig.height=6}
siteNumberSP <- c("11113500", "11118500")
parameterSP <- "00060" #Discharge 
startDateSP <- "2021-10-04"
endDate <- as.Date("2021-10-06")

both_info <- readNWISdata(sites = siteNumberSP, 
                          service = "iv", 
                          parameterCd = parameterSP, 
                          startDate = startDateSP, 
                          endDate = endDate) %>% 
  rename(discharge = X_00060_00000) %>% 
  separate(dateTime, c("date", "time"), sep = " ", remove = FALSE) %>% 
  filter(date < "2021-10-07")

ggplot(data = both_info, aes(x = dateTime, y = discharge)) +
  geom_line(aes(color = site_no)) +
  labs(y = "Discharge (cubic feet)", 
       x = "Date", 
       title = "Ventura River and Santa Paula Creek Discharge (Oct 4 - 6 2021)") +
  scale_color_manual(name = "Site ID", 
                     labels = c("Santa Paula", "Ventura"), 
                     values = c("cyan3", "red")) +
  scale_x_datetime(breaks = date_breaks("2 hour"), 
                   labels = date_format(format = "%m-%d %H:%m")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
        panel.background = element_rect("white"), 
        axis.line = element_line(size = 0.5, 
                                 linetype = "solid",
                                 colour = "black"), 
        panel.grid.major.x = element_line(size = 0.5, 
                                        linetype = 'solid',
                                        colour = "grey"),
        panel.grid.minor.x = element_line(size = 0.25, 
                                        linetype = 'solid',
                                        colour = "lightgrey")
        ) 

```

The best way to query multiple datasets using this package is to pass a vector of site numbers as the first argument to `readNWISdata()`. You can then wrangle the data however you see fit after it is all read in, filtering based on the `siteno`. 

The stream flow in the Santa Paula Creek peaked at 11:10 on October 5th. There was no significant peak in the Ventura River discharge, however the entire day on October 5th had higher discharge than the day before and after. There was likely another precipitation event affecting the peak seen on October 6th for both Santa Paula and Ventura's discharges. 


# Assignment 2.2: Using metajam

```{r, message=FALSE}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"
#metajam::download_d1_data(data_url = data_url, here::here("data"))
data <- read_d1_files(here::here("data", "household_language_files"))
household_data <- data$data
only_english <- household_data %>% 
  clean_names() %>% 
  filter(year >= 2009, 
         year <= 2015, 
         !total == 0) %>% 
  select(year, sasap_region, city, total, speak_only_english) %>% 
  group_by(year) %>% 
  summarize(yearly_percent_avg=mean(speak_only_english/total))

ggplot(only_english, aes(x = year, y = yearly_percent_avg)) +
  geom_line() +
  labs(x = "Year", 
       y = "Yearly Average Percentage", 
       title = "Average Percentage of Alaskan Households Speaking English") +
  theme_classic()
```

